### 2020-08-20 paper/theory/model_executive_bargain.tex

\begin{document}

@@

In this section, I adapt the canonical vote-buying model proposed by Groseclose and Snyder (1996) and Banks (2000) to model how executive control over the city council determines patronage. The theoretical framework centers on two main actors, a government (\emph{governo}) and opposition (\emph{oposição}) who compete for votes in the city council.\footnote{I use the terms mayor and government interchangeably for the rest of this section.} The mayor seeks to enact her preferred policy, but can only do so by passing a simple majority vote in the council. The opposition, on the other hand, wants to block her efforts and maintain the status quo. In order to build a winning coalition, the mayor must spend a key political resource: patronage appointments into the municipal bureaucracy.

Ultimately, two forces shape the mayor's strategy: the amount of resources controlled by the opposition and the underlying preferences of the legislators. The key parameter of interest is the underlying preferences -- i.e. support for the mayor -- of city councilors. In some municipalities a mayor has widespread political support ("base política"), while in others, the opposition ("oposição") garners more seats. In the actual empirical estimation, we approximate these preferences by looking into the councilor's partisan alignments in the mayoral election. For the second determinant, the opposition's resource, it is clear that an opposition with more resources can better compete with the mayor, forcing the latter to spend more to deter them.

I show that, in equilibrium, the cost of passing the mayor's preferred policy decreases monotonically as the legislative support for the mayor increases. Substantively, the model predicts that municipalities in which the mayor has more (less) councilors favorable to her, we should observe less (more) patronage. In the remainder of this section, I outline the setting of the model and derive key comparative statics that guide the empirical estimation of this paper. I conclude with a discussion of the model and how it applies to the Brazilian municipal context.

@@

\end{document}

>>>>

\bibliographystyle{plain}
\begin{document}

@@

In this section, I use the canonical vote-buying model proposed by \citet{groseclose_1996_buying} and \citet{banks_2000_buying} to model mayor's strategy to engage in patronage in Brazil. For any given term, a mayor seeks to enact her preferred policy, but to do so needs to garner legislative support in the city council. The preferences of these legislators vary: some of these councilors are more inclined to supporting the government (\emph{governista}), while others would rather oppose it. These different preferences translate into differential costs to sway their vote: city councilors aligned with the mayor are cheaper than their opposition counterparts.

In this set-up, the mayor competes with the opposition to buy votes. The strategy of the mayor focuses on how to pre-emptively coopt some of these councilors so that they ultimately vote in favor of the mayor's proposal. Building this winning coalition requires side payments of various forms, but my primary focus is on patronage appointments. City councilors who vote for the mayor's proposal are rewarded, as loyalists, with positions in government. This arrangement is mutually beneificial: as pointed out by \citet{robinson_political_2013}, these appointments tie the councilor's fate to the patron. Because patronage appointments are discretionary, in order to maintain these appointments the city councilor must remain loyal to the government and vote in favor of the mayor's proposal.

The main prediction of the model is that, in equilibrium, the cost of passing the mayor's preferred policy decreases monotonically as the legislative support for the mayor increases. Substantively, the model predicts that municipalities in which the mayor has more (less) councilors favorable to her, we should observe less (more) patronage. Fragmented governments, marked by a strong opposition in the legislature, therefore translate into a greater degree of patronage, with negative consequences for the electorate. In the remainder of this section, I outline the setting of the model and derive key comparative statics that guide the empirical estimation of this paper. I conclude with a discussion of the model and how it applies to the Brazilian municipal context.

@@

\bibliography{../paper/dissertation}
\end{document}
